Rhizaria in the oligotrophic ocean exhibit clear temporal and vertical variability.

### 2024-07-06 manuscript/main.tex
minor edit and render

\usepackage{amsmath,amssymb}
\usepackage{setspace}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[left=1in,textwidth=6.5in,right=1in]{geometry}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\makeatletter
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}{
    \@ifstar
      \xxxParagraphStar
      \xxxParagraphNoStar
  }
  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}{
    \@ifstar
      \xxxSubParagraphStar
      \xxxSubParagraphNoStar
  }
  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\makeatother

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

@@

% Add any tex header commands here
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Rhizaria in the oligotrophic ocean exhibit clear temporal and vertical variability.},
  pdfauthor={Alex Barth*; Leocadio Blanco-Berical; Rod Johnson; Joshua Stone},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\end{minipage}  \end{titlepage}

\setstretch{2}
\section{Introduction}\label{introduction}

@@

\section{Methods}\label{methods}

\subsection{Oceanographic sampling}\label{oceanographic-sampling}

@@

\subsection{Rhizaria imaging processing and abundance
quantification}\label{rhizaria-imaging-processing-and-abundance-quantification}

@@

\begin{figure}[H]

@@

\end{figure}%

The UVP5 samples at \textasciitilde15Hz rate as it descends the water
column and records the exact position of each particle larger than
600\(\mu m\) (ESD). However, identified rhizaria ranged from a
934\(\mu m\) ESD Aulacanthidae cell to a Collodarian colony over 10mm in
diameter. To confirm that the UVP5 was sampling adequately across all
size ranges, an normalized biomass size spectrum (NBSS) slope was
constructed to identify a drop-off which would indicate poor-sampling at
the small size range (Lombard \emph{et al.}, 2019; Barth and Stone,
2024). However, it was evident from this analysis that all size ranges
were adequately sampled across the size range (Supplemental Figure 2) so
no data were excluded. The UVP5 reports the exact depth at which a
particle is recorded, however to estimate abundance, observations must
be binned over fixed depth intervals. Our deployments had variable
descent depths and speeds with more casts descending to 500m than 1000m
and descents quicker through the epipelagic than the mesopelagic (see
Barth and Stone (2022) for an extended discussion of UVP5 data
processing). For the present study, Rhizaria abundances were estimated
in 25m vertical bins, which offer a moderate sampling volume per bin
(average 0.948\(m^3\) in the epipelagic and 0.589\(m^3\) in the
mesopelagic; Supplemental Table 1) while still maintaining ecologically
relevant widths. However, concentrations in a 25m bin would need to be
greater than 2.428 ind. \(m^{-3}\) and 3.912 ind. \(m^{-3}\), in the
epipelagic and mesopelagic respectively, to fall below a 10\%
non-detection risk (Benfield \emph{et al.}, 1996; Barth and Stone,
2024). Because we typically observed many Rhizaria taxa below these
concentrations, we present the 25m binned data to visualize broad-scale
average distributions. For quantifying and modelling Rhizaria
abundances, we present integrated abundance estimates, with each cast.
Due to the variable descent depths of the UVP, data are categorized as
epipelagic (0-200m), upper mesopelagic (200-500m), and lower mesopelagic
(500-1000m). The average sampling volume integrated through these
regions were 7.59\(m^2\), 7.06\(m^2\), and 11.77\(m^2\), with
non-detection thresholds at 0.30 ind. \(m^{-2}\), 0.33 ind. \(m^{-2}\),
and 0.20 ind. \(m^{-2}\) respectively. All UVP data processing was done
using the \texttt{EcotaxaTools} package in R (Barth 2023).

\subsection{Modelling environmental controls of Rhizaria
abundance}\label{modelling-environmental-controls-of-rhizaria-abundance}

@@

\section{Results}\label{results}

\subsubsection{Environmental
Variability}\label{environmental-variability}

@@

\end{figure}%

@@

\subsubsection{Rhizaria abundance and
distribution}\label{rhizaria-abundance-and-distribution}

@@

\subsubsection{Body size throughout the water
column.}\label{body-size-throughout-the-water-column.}

@@

\end{figure}%

\subsubsection{Environmental Drivers of Rhizaria
Abundance}\label{environmental-drivers-of-rhizaria-abundance}

\begin{figure}[H]

@@

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.4300}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.3400}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.0800}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.0800}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.0800}}@{}}
\caption{Generalized Additive Model results for integrated total
Rhizaria abundance in different regions of the water column. edf =
estimated degrees of freedom, F = F-statistic, p = p-value. Models were
selected using a backwards-stepwise approach with a retention threshold
of p = 0.1 or a large enough reduction in model fit.}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
\textbf{Model}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Term
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
edf
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
F
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
p
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
\textbf{Model}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Term
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
edf
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
F
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
p
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{All Rhizaria Epipelagic} & Temperature & 0.7950 & 0.492 &
0.0440 \\
\(R^2_{adj}\)= 0.48 & Salinity & 2.4117 & 5.113 & \textless0.001 \\
& Silica & 2.1511 & 8.601 & \textless0.001 \\
& Bacteria Concentration & 0.8870 & 1.544 & \textless0.001 \\
& Avg Mass Flux & 1.4107 & 0.802 & 0.0211 \\
& Primary Productivity & 1.9280 & 2.736 & \textless0.001 \\
& Particle Concentration & 0.9151 & 2.136 & \textless0.001 \\
\textbf{All Rhizaria Upper Mesopelagic} & Silica & 0.6702 & 0.404 &
0.0814 \\
\(R^2_{adj}\)= 0.40 & Avg N Flux & 2.9524 & 3.172 & \textless0.001 \\
& Particle Concentration & 3.5568 & 11.82 & \textless0.001 \\
\textbf{All Rhizaria Lower Mesopelagic} & Avg Mass Flux & 2.036 & 2.953
& \textless0.001 \\
\(R^2_{adj}\)= 0.61 & Particle Concentration & 3.819 & 27.43 &
\textless0.001 \\
\end{longtable}

@@

\toprule\noalign{}

@@

\midrule\noalign{}

@@

\bottomrule\noalign{}
\endlastfoot

@@

@@

\section{Discussion}\label{discussion}

\subsubsection{Overall Rhizaria abundance and
patterns}\label{overall-rhizaria-abundance-and-patterns}

@@

\subsubsection{Relationship to environmental
parameters}\label{relationship-to-environmental-parameters}

@@

\subsubsection{Vertical Structure and Trophic
Roles}\label{vertical-structure-and-trophic-roles}

In this study we present a clear pattern of vertical zonation between
different Rhizaria groups. Largely, the taxonomic composition and
vertical positioning were similar to Rhizaria zonation in the California
Current Ecosystem (Biard and Ohman, 2020). It should be noted however,
that the secondary abundance peak reported in the present study is
lower. This is likely due to the more oligotrophic nature of the study
region, were the euphotic zone penetrates deeper into the water column.
Most prevalent in the epipelagic were Collodaria. These mixotrophic
Radiolaria have long been reported to contribute to primary productivity
in the euphotic zone (Michaels \emph{et al.}, 1995; Dennett \emph{et
al.}, 2002). Collodaria are thought to be particularly successful
globally in oligotrophic regions due to their photosymbiotic
relationships (Biard \emph{et al.}, 2016, 2017). We observed the highest
abundance of Collodaria during June 2021, supporting the notion they can
thrive during the typically low-nutrient conditions of summer
stratification. However, Collodaria also increased during the spring
mixing period, suggesting that they can thrive during conditions which
may typically be thought to favor autotrophs. Furthermore, while
Collodaria were primarily absent from below 250m, there were a few
instances of deeper colonies being observed. Global investigations of
polycystine flux, suggest that deep-Collodaria in oligotrophic regions
may be a consequence of isothermal submersion (Boltovskoy, 2017).
Alternatively, surface waters at BATS often mix into the mode water
during the seasonal mixing, so Collodaria in the deeper waters may be a
result of diapyncal mixing. Another effectively exclusively epipelagic
Rhizaria was the Phaeodaria family of Castanellidae. All Phaeodaria are
thought to be fully heterotrophic (Nakamura and Suzuki, 2015),
nonetheless a number of studies, including this one, report
Castanellidae to be typically found in the lower epipelagic (Zasko and
Rusanov, 2005; Biard \emph{et al.}, 2018; Biard and Ohman, 2020). It
should be considered that perhaps Castanellidae specializes in feeding
on sinking particles directly at the base of the epipelagic. Given it's
smaller size (Nakamura and Suzuki, 2015), Castanellidae does not need a
large diameter to efficiently flux feed at the typically particle rich
region of the lower epipelagic.

>>>>

\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{setspace}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[left=1in,textwidth=6.5in,right=1in]{geometry}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

@@

% Add any tex header commands here
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Rhizaria in the oligotrophic ocean exhibit clear temporal and vertical variability.},
  pdfauthor={Alex Barth*; Leocadio Blanco-Berical; Rod Johnson; Joshua Stone},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\end{minipage}  \end{titlepage}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, interior hidden, sharp corners, borderline west={3pt}{0pt}{shadecolor}, frame hidden, enhanced, boxrule=0pt]}{\end{tcolorbox}}\fi

\setstretch{2}
\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

@@

\hypertarget{methods}{%
\section{Methods}\label{methods}}

\hypertarget{oceanographic-sampling}{%
\subsection{Oceanographic sampling}\label{oceanographic-sampling}}

@@

\hypertarget{rhizaria-imaging-processing-and-abundance-quantification}{%
\subsection{Rhizaria imaging processing and abundance
quantification}\label{rhizaria-imaging-processing-and-abundance-quantification}}

@@

\begin{figure}

@@

\end{figure}

The UVP5 samples at \textasciitilde15Hz rate as it descends the water
column and records the exact position of each particle larger than
600\(\mu m\) (ESD). However, identified rhizaria ranged from a
934\(\mu m\) ESD Aulacanthidae cell to a Collodarian colony over 10mm in
diameter. To confirm that the UVP5 was sampling adequately across all
size ranges, an normalized biomass size spectrum (NBSS) slope was
constructed to identify a drop-off which would indicate poor-sampling at
the smaller end of the size range (Lombard \emph{et al.}, 2019; Barth
and Stone, 2024). However, it was evident from this analysis that all
size ranges were adequately sampled across the size range (Supplemental
Figure 2) so no data were excluded. The UVP5 reports the exact depth at
which a particle is recorded, however to estimate abundance,
observations must be binned over fixed depth intervals. Our deployments
had variable descent depths and speeds with more casts descending to
500m than 1000m and descents quicker through the epipelagic than the
mesopelagic (see Barth and Stone (2022) for an extended discussion of
UVP5 data processing). For the present study, Rhizaria abundances were
estimated in 25m vertical bins, which offer a moderate sampling volume
per bin (average 0.948\(m^3\) in the epipelagic and 0.589\(m^3\) in the
mesopelagic; Supplemental Table 1) while still maintaining ecologically
relevant widths. However, concentrations in a 25m bin would need to be
greater than 2.428 ind. \(m^{-3}\) and 3.912 ind. \(m^{-3}\), in the
epipelagic and mesopelagic respectively, to fall below a 10\%
non-detection risk (Benfield \emph{et al.}, 1996; Barth and Stone,
2024). Because we typically observed many Rhizaria taxa below these
concentrations, we present the 25m binned data to visualize broad-scale
average distributions. For quantifying and modelling Rhizaria
abundances, we present integrated abundance estimates, with each cast.
Due to the variable descent depths of the UVP, data are categorized as
epipelagic (0-200m), upper mesopelagic (200-500m), and lower mesopelagic
(500-1000m). The average sampling volume integrated through these
regions were 7.59\(m^2\), 7.06\(m^2\), and 11.77\(m^2\), with
non-detection thresholds at 0.30 ind. \(m^{-2}\), 0.33 ind. \(m^{-2}\),
and 0.20 ind. \(m^{-2}\) respectively. All UVP data processing was done
using the \texttt{EcotaxaTools} package in R
(\url{https://github.com/TheAlexBarth/EcotaxaTools?tab=readme-ov-file}).

\hypertarget{modelling-environmental-controls-of-rhizaria-abundance}{%
\subsection{Modelling environmental controls of Rhizaria
abundance}\label{modelling-environmental-controls-of-rhizaria-abundance}}

@@

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{environmental-variability}{%
\subsubsection{Environmental
Variability}\label{environmental-variability}}

@@

\end{figure}

@@

\hypertarget{rhizaria-abundance-and-distribution}{%
\subsubsection{Rhizaria abundance and
distribution}\label{rhizaria-abundance-and-distribution}}

@@

\hypertarget{body-size-throughout-the-water-column.}{%
\subsubsection{Body size throughout the water
column.}\label{body-size-throughout-the-water-column.}}

@@

\end{figure}

\hypertarget{environmental-drivers-of-rhizaria-abundance}{%
\subsubsection{Environmental Drivers of Rhizaria
Abundance}\label{environmental-drivers-of-rhizaria-abundance}}

\begin{figure}

@@

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.4300}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.3400}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.0800}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.0800}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.0800}}@{}}
\caption{Generalized Additive Model results for integrated total
Rhizaria abundance in different regions of the water column. edf =
estimated degrees of freedom, F = F-statistic, p = p-value. Models were
selected using a backwards-stepwise approach with a retention threshold
of p = 0.1 or a large enough reduction in model fit.}\tabularnewline
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
\textbf{Model}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Term
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
edf
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
F
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
p
\end{minipage} \\
\midrule()
\endfirsthead
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
\textbf{Model}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Term
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
edf
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
F
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
p
\end{minipage} \\
\midrule()
\endhead
\textbf{All Rhizaria Epipelagic} & Temperature & 0.7950 & 0.492 &
0.0440 \\
\(R^2_{adj}\)= 0.48 & Salinity & 2.4117 & 5.113 & \textless0.001 \\
& Silica & 2.1511 & 8.601 & \textless0.001 \\
& Bacteria Concentration & 0.8870 & 1.544 & \textless0.001 \\
& Avg Mass Flux & 1.4107 & 0.802 & 0.0211 \\
& Primary Productivity & 1.9280 & 2.736 & \textless0.001 \\
& Particle Concentration & 0.9151 & 2.136 & \textless0.001 \\
\textbf{All Rhizaria Upper Mesopelagic} & Silica & 0.6702 & 0.404 &
0.0814 \\
\(R^2_{adj}\)= 0.40 & Avg N Flux & 2.9524 & 3.172 & \textless0.001 \\
& Particle Concentration & 3.5568 & 11.82 & \textless0.001 \\
\textbf{All Rhizaria Lower Mesopelagic} & Avg Mass Flux & 2.036 & 2.953
& \textless0.001 \\
\(R^2_{adj}\)= 0.61 & Particle Concentration & 3.819 & 27.43 &
\textless0.001 \\
\bottomrule()
\end{longtable}

@@

\toprule()

@@

\midrule()

@@

@@

\bottomrule()

@@

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

\hypertarget{overall-rhizaria-abundance-and-patterns}{%
\subsubsection{Overall Rhizaria abundance and
patterns}\label{overall-rhizaria-abundance-and-patterns}}

@@

\hypertarget{relationship-to-environmental-parameters}{%
\subsubsection{Relationship to environmental
parameters}\label{relationship-to-environmental-parameters}}

@@

\hypertarget{vertical-structure-and-trophic-roles}{%
\subsubsection{Vertical Structure and Trophic
Roles}\label{vertical-structure-and-trophic-roles}}

In this study we present a clear pattern of vertical zonation between
different Rhizaria groups. Largely, the taxonomic composition and
vertical positioning were similar to Rhizaria zonation in the California
Current Ecosystem (Biard and Ohman, 2020). It should be noted however,
that the secondary abundance peak reported in the present study is lower
in magnitude. This is likely due to the more oligotrophic nature of the
study region, were the euphotic zone penetrates deeper into the water
column. Most prevalent in the epipelagic were Collodaria. These
mixotrophic Radiolaria have long been reported to contribute to primary
productivity in the euphotic zone (Michaels \emph{et al.}, 1995; Dennett
\emph{et al.}, 2002). Collodaria are thought to be particularly
successful globally in oligotrophic regions due to their photosymbiotic
relationships (Biard \emph{et al.}, 2016, 2017). We observed the highest
abundance of Collodaria during June 2021, supporting the notion they can
thrive during the typically low-nutrient conditions of summer
stratification. However, Collodaria also increased during the spring
mixing period, suggesting that they can thrive during conditions which
may typically be thought to favor autotrophs. Furthermore, while
Collodaria were primarily absent from below 250m, there were a few
instances of deeper colonies being observed. Global investigations of
polycystine flux, suggest that deep-Collodaria in oligotrophic regions
may be a consequence of isothermal submersion (Boltovskoy, 2017).
Alternatively, surface waters at BATS often mix into the mode water
during the seasonal mixing, so Collodaria in the deeper waters may be a
result of diapyncal mixing. Another effectively exclusively epipelagic
Rhizaria was the Phaeodaria family of Castanellidae. All Phaeodaria are
thought to be fully heterotrophic (Nakamura and Suzuki, 2015),
nonetheless a number of studies, including this one, report
Castanellidae to be typically found in the lower epipelagic (Zasko and
Rusanov, 2005; Biard \emph{et al.}, 2018; Biard and Ohman, 2020). It
should be considered that perhaps Castanellidae specializes in feeding
on sinking particles directly at the base of the epipelagic. Given it's
smaller size (Nakamura and Suzuki, 2015), Castanellidae does not need a
large diameter to efficiently flux feed at the typically particle rich
region of the lower epipelagic.

The mesopelagic generally was home to known heterotrophic organisms,
particularly for those which were constrained to exclusively occupy
deeper waters. This is consistent with Blanco-Bercial \emph{et al.}
(2022)'s observation of an auto-/mixotroph to heterotroph gradient in
local protist community as well as global patterns in Rhizaria ecology
(Laget \emph{et al.}, 2024). The upper mesopelagic interestingly had
relatively low total abundance. This low-abundance region likely
reflects the dynamics of productivity and export throughout the water
column. While productivity and thus sinking particles for flux feeders
are high in the euphotic zone, much of this is attenuated throughout the
epipelagic. So, while the base of the epipelagic may provide a rich
feeding environment for Castanellidae, smaller protists, or
heterotrophic bacteria (Figure 2F), the region from 200-500m might be
otherwise food poor. Perhaps it is more advantageous for Rhizaria to
situate deeper, in darker regions of the twilight zone where visual
predators and vertically migrating organisms may not feed on them. Also
it should be noted that Phaeodaria utilize silica to build their opaline
tests, and silica concentrations started to increase around 500m (Figure
2G). However, \(Si\) was not a significant term for any taxon-specific
model in the mesopelagic. Aulosphaeridae was only found to have
significant relationships, although weak fits, to particle concentration
in the mesopelagic. In our study, while consistently observed, overall
abundances of Aulosphaeridae were very low. In the Pacific Ocean, on
California's Coast, much higher abundances of Aulosphaeridae have been
reported (Zasko and Rusanov, 2005; Biard and Ohman, 2020) and they have
massive potential to impact silica export (Biard \emph{et al.}, 2018).
Coelodendridae were also seemingly restricted to the deeper section of
the mesopelagic. This is interesting given that in the California
Current, (Biard and Ohman, 2020) found a bimodal distribution in
Coelodendridae. There are several morphotypes corresponding to different
taxa of Coelodendridae (Nakamura and Suzuki, 2015; Biard and Ohman,
2020). So it may be that only a few types of Coelodendridae were
observed in this study, while the epipelagic variety was not.
Alternatively, the lower epipelagic of the extremely productive
California Current may provide adequate habitat for Coelodendridae,
which is not available in the oligotrophic Sargasso Sea.
